---
title: "Climate_change_proj"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(openxlsx)
library(tibble)
library(forcats)
library(scales)
library(viridis)
library(cols4all)
library(maps)
library(mapproj)
library(ftplottools)
library(shiny)
```


# Who rules the world? Oil

## Climate change, a study of sources and consequences

Introduction:

The aim of this piece is to detail the energy sources in the European countries, while also looking at how much emissions are generated. Furthermore, in an ambitious attempt to contextualize it the background will be enhanced with a detail of the production and consumption of energy by different sectors and industries.

"order the data do more natural don't cite as in usual"

```{r Load data, include=FALSE, message=FALSE}
#Energy/ electricity production
data_supply_raw<-read.csv("Data/data_en_supp.csv")%>% select(-c(freq,unit,nrg_bal,OBS_FLAG,DATAFLOW,LAST.UPDATE)) %>%mutate_at(., "OBS_VALUE", ~replace(., is.na(.), 0))%>% mutate (OBS_VALUE=OBS_VALUE/85.985*1000)

data_prod_raw<-read.csv("Data/data_en_cons.csv") %>% select(-c(freq,unit,nrg_bal,OBS_FLAG,DATAFLOW,LAST.UPDATE)) %>%  mutate_at(., "OBS_VALUE", ~replace(., is.na(.), 0)) %>% mutate (OBS_VALUE=OBS_VALUE/85.985*1000)

elec_prod<-read.csv("Data/energy_prod_monthly_europe.csv") %>% select(-c(freq,unit,OBS_FLAG,DATAFLOW,LAST.UPDATE))%>%filter(siec!="RA410",siec!="RA420",siec!="RA320",siec!="RA310",siec!="RA110",siec!="RA120",siec!="RA130")

# Division by sectors and performances:

dictionary_countries <- data.frame(readRDS("Data/country_codenames.rds"))
colnames(dictionary_countries)<- c("Country_codes","Country")

dictionary_countries <- dictionary_countries %>% mutate(Country=ifelse(Country=="United Kingdom (Convention)","United Kingdom",Country),Country_codes=ifelse(Country_codes=="GR","EL",Country_codes))


#Emissions by country:
eur_em<-read.xlsx("Data/EU_co2_emiss.xlsx")
eur_em_predict<-read.csv("Data/EU_co2_emiss_proj.csv") %>% select(-c(Reported.Value,RY.calibration,SubmissionYear))

selection_countries<- dictionary_countries %>% filter(Country != "United Kingdom (Convention)",Country !="EU (KP)",Country !="EU (KP)",Country !="EU-27",Country!="EU-27+UK")


#Energy prod broken down by products and countries:
# w_sec_count_em<- read.xlsx("Data/Energy_production_country_sector.xlsx") %>% filter(Country %in% selection_countries$Country)
# saveRDS(w_sec_count_em,"Data/energy_sector_country.rds")
w_sec_count_em<-readRDS("Data/energy_sector_country.rds")
## There's also the investment file which is to be used on a more discretionary manner not as a whole don't open but it's the full file:
#poll<-read.csv("Data/pollution_by_country.csv") --- not really needed anymore.
```


```{r dictionaries, include=FALSE}
dic_2<-data.frame(siec = unique(elec_prod $siec),Fuel=c("Coal & gases","Combustible fuels","Combustible fuels","Biofuels","Natural gas","Nuclear","Oil & petrol","Hydro",
                                                   "Geothermal","Wind","Solar","Other renewables","Total","Other"))
elec_broken<-left_join(dic_2,elec_prod,by="siec") %>% select(-c(siec))

dic_3<-data.frame(siec = unique(elec_prod $siec),Fuel=c("Coal & gases","Combustible fuels","Combustible fuels","Biofuels","Natural gas","Nuclear","Oil & petrol","Renewables","Renewables",
                                                   "Renewables","Renewables","Renewables","Total","Other"))
elec<-left_join(dic_3,elec_prod,by="siec") %>% select(-c(siec))

dictionary<-data.frame(siec=unique(data_prod_raw$siec),Fuel=c("Coal & gases","Coal & gases","Electricity","Natural gas","Heat","Nuclear","Oil & petrol","Peat","Renewables/biofuel","Oil & shale","Total","Non-ren waste"))

data_prod<-left_join(dictionary,data_prod_raw,by="siec")%>% select(-c(siec))


data_prod$TIME_PERIOD<-as.Date(ISOdate(data_prod$TIME_PERIOD, 1, 1)) 
elec$TIME_PERIOD<-as.Date(paste(elec$TIME_PERIOD, 1,sep="-"))
elec_broken$TIME_PERIOD<-as.Date(paste(elec$TIME_PERIOD, 1,sep="-"))

# data_prod %>% filter(Fuel =="Oil & shale",OBS_VALUE!=0) %>% select(geo,TIME_PERIOD,OBS_VALUE) 

#for elecctricy combine renewables-biofuel & combustible fuels bag into coal & gases

data_prod$Fuel[data_prod$Fuel=="Oil & shale"]="Oil & petrol"

elec <- elec %>% mutate(
  Fuel = ifelse(Fuel == "Combustible fuels","Coal & gases",Fuel ),
  Fuel = ifelse(Fuel == "Renewables" | Fuel == "Biofuels" , "Renewables/biofuel",Fuel )
)


```


```{r data available heads,include=FALSE,echo=FALSE, message=FALSE}
head(data_prod_raw)
head(data_supply_raw)
# From the EEA comission 

head(eur_em)
head(eur_em_predict)

head(w_sec_count_em)
```

# 1- Energy production in Europe - which type is the most used/which is the trend?


## 1.1 - Plots of production in Europe
```{r plots country,include=FALSE,warning=FALSE, message=FALSE}
#Totals by country, choose whether to breakdown electricity or not:

country_totals_byfuel<- data_prod %>% filter(Fuel != "Total", geo== "EU27_2020",geo!= "EA19")

total_country_totals <- country_totals_byfuel %>% 
  group_by(TIME_PERIOD,Fuel) %>%
  summarise(Consumed = sum(OBS_VALUE))

ggplot(total_country_totals) + aes(x=TIME_PERIOD, y=Consumed,col =Fuel) + geom_line()


```



```{r warning=FALSE, message=FALSE, echo=FALSE}
## Combination elec + productions and plotting:

elec_years <- elec %>% filter(Fuel!= "Total") %>%
  mutate( year = format(TIME_PERIOD, "%Y")) %>%
  group_by(year,Fuel,geo) %>% 
  summarise(OBS_VALUE = sum(OBS_VALUE)) %>%
  mutate(TIME_PERIOD = as.Date(paste(year,1,1,sep="-"))) %>% ungroup(year) %>% select(-year)

elec_plus_production<- data.frame(rbind(data_prod%>% filter(Fuel!="Electricity",Fuel!="Total"), elec_years )) %>% filter(geo== "EU27_2020",TIME_PERIOD>= "2017-01-01",TIME_PERIOD<"2021-01-01") %>% select(-geo)

broken_ele_prod <- elec_plus_production %>% 
  group_by(TIME_PERIOD,Fuel) %>%
  summarise(Consumed = sum(OBS_VALUE))

prop_elec_2020<-broken_ele_prod %>% filter(TIME_PERIOD =="2020-01-01") %>%
  arrange(desc(Consumed)) %>%  
  mutate( Fuel=factor(Fuel,levels=Fuel),
          Fuel = fct_collapse(Fuel,"Other"=c("Non-ren waste","Other","Peat","Heat")))%>%
  group_by(TIME_PERIOD,Fuel) %>%
  summarise(Consumed = sum(Consumed)) %>% mutate(prop= Consumed/sum(Consumed))

prop_elec_2017<-broken_ele_prod %>% filter(TIME_PERIOD =="2017-01-01") %>%
  arrange(desc(Consumed)) %>%  
  mutate( Fuel=factor(Fuel,levels=Fuel),
          Fuel = fct_collapse(Fuel,"Other"=c("Non-ren waste","Other","Peat","Heat")))%>%
  group_by(TIME_PERIOD,Fuel) %>%
  summarise(Consumed = sum(Consumed)) %>% mutate(prop= Consumed/sum(Consumed))

preserving_orders<-merge(prop_elec_2020,prop_elec_2017,by="Fuel") %>% arrange(desc(prop.y))

elec_prod_clean<-broken_ele_prod %>% 
  arrange(desc(Consumed)) %>%  
  mutate( Fuel=factor(Fuel,levels=Fuel),
          Fuel = fct_collapse(Fuel,"Other"=c("Non-ren waste","Other","Peat","Heat")))%>%
  group_by(TIME_PERIOD,Fuel) %>%
  summarise(Consumed = sum(Consumed)) %>% ungroup(TIME_PERIOD,Fuel)


### Electricity and energy consumption

plot<-ggplot(elec_prod_clean,aes(x=TIME_PERIOD, y=Consumed/10**6, fill=Fuel)) + 
            geom_area(alpha=0.6 , size=1, colour="black") +
            labs(x="Year",
                 y=expression(paste('Energy consumed (',10^6,' GWE)')),
                 title="Cummulative energy production (breaking down electricity)") + 
            theme_bw() +
            scale_fill_brewer(palette = "RdYlBu")

plot + annotate("label",
                x = c(rep(as.Date("2020-01-01","%Y-%m-%d"),6)),
                y = c(9.8,6.5,4.5,2.5,1,0.2),
                label = paste(round(preserving_orders$prop.x,3)*100,"%"), 
                color="black", 
                size=2.5 , 
                fontface="bold") +
  annotate("label",
           x = c(rep(as.Date("2017-01-01","%Y-%m-%d"),6)),
           y = c(10.2,6.8,4.7,2.2,1,0.2),
           label = paste(sort(round(prop_elec_2017$prop,3)*100,decreasing=TRUE),"%"), 
           color="black", 
           size=2.5 , 
           fontface="bold")
```



```{r With electricity as factor, warning=FALSE, message=FALSE, echo= FALSE}
energy_w_elec<- data.frame(data_prod) %>% 
  filter(geo== "EU27_2020",Fuel!="Total") %>% 
  select(-geo)

energy_w_elec <- energy_w_elec %>% 
  group_by(TIME_PERIOD,Fuel) %>%
  summarise(Consumed = sum(OBS_VALUE))

energy_at_2020 <- energy_w_elec %>% filter(TIME_PERIOD =="2020-01-01") %>%
  arrange(desc(Consumed)) %>%  
  mutate( Fuel=factor(Fuel,levels=Fuel),
          Fuel = fct_collapse(Fuel,"Other"=c("Non-ren waste","Other","Peat","Heat")))%>%
  group_by(TIME_PERIOD,Fuel) %>%
  summarise(Consumed = sum(Consumed)) %>% mutate(prop= Consumed/sum(Consumed))

energy_at_2017 <- energy_w_elec %>% filter(TIME_PERIOD =="2017-01-01") %>%
  arrange(desc(Consumed)) %>%  
  mutate( Fuel=factor(Fuel,levels=Fuel),
          Fuel = fct_collapse(Fuel,"Other"=c("Non-ren waste","Other","Peat","Heat")))%>%
  group_by(TIME_PERIOD,Fuel) %>%
  summarise(Consumed = sum(Consumed)) %>% mutate(prop= Consumed/sum(Consumed))

preserving_orders<-merge(energy_at_2020,energy_at_2017,by="Fuel") %>% arrange(desc(prop.y))

prod_clean<-energy_w_elec %>% 
  arrange(desc(Consumed)) %>%  
  mutate( Fuel=factor(Fuel,levels=Fuel),
          Fuel = fct_collapse(Fuel,"Other"=c("Non-ren waste","Other","Peat","Heat")))%>%
  group_by(TIME_PERIOD,Fuel) %>%
  summarise(Consumed = sum(Consumed)) %>% ungroup(TIME_PERIOD,Fuel)

### Energy consumption + electricity: 

plot_2<-ggplot(prod_clean,aes(x=TIME_PERIOD, y=Consumed/10**6, fill=Fuel)) + 
              geom_area(alpha=0.6 , size=1, colour="black") + 
            labs(x="Year",
                 y=expression(paste('Energy consumed (',10^6,' GWE)')),
                 title="Cummulative energy production inc. electricity") + 
            theme_bw() +
            scale_fill_brewer(palette = "RdYlBu") + 
            theme(axis.text.x = element_text(angle = 45,hjust=1),panel.grid = element_blank())

plot_2 + annotate("label",x = c(rep(as.Date("2020-01-01","%Y-%m-%d"),6)),
                   y = c(9,5.2,3,1.5,0.6,0),
                   label = paste(round(preserving_orders$prop.x,3)*100,"%"), color="black", 
                    size=2.5 , fontface="bold") +
  annotate("label",x = c(rep(as.Date("2011-01-01","%Y-%m-%d"),6)),
                   y = c(9,5.2,3,1.5,0.6,0),
                   label = paste(sort(round(preserving_orders$prop.y,3)*100,decreasing=TRUE),"%"), color="black", 
                   size=2.5 , fontface="bold")
```

```{r Electricity as factor with shiny}
countries <- dictionary_countries %>% 
            transmute(Country=ifelse(Country=="United Kingdom (Convention)","United 
                                  Kingdom",Country),
                      geo=ifelse(Country_codes=="GE","EL",Country_codes),
                      geo=ifelse(Country_codes=="EUX","EU27_2020",Country_codes))

european_data<-right_join(data_prod,countries,by="geo")

server <- function(input, output) {
  
  prod_clean_shiny<- reactive({data.frame(european_data) %>% 
  filter(Country== input$region,!Fuel%in%c("Total","Nuclear")) %>% 
  select(-c(geo,Country))%>%
  group_by(TIME_PERIOD,Fuel) %>%
  summarise(Consumed = sum(OBS_VALUE))%>%
      arrange(desc(Consumed)) %>%  
  mutate( Fuel=factor(Fuel,levels=Fuel),
          Fuel = fct_collapse(Fuel,"Other"=c("Non-ren waste","Other","Peat","Heat")))%>%
  group_by(TIME_PERIOD,Fuel) %>%
  summarise(Consumed = sum(Consumed)) %>% ungroup(TIME_PERIOD,Fuel)
    })

  # Fill in the spot we created for a plot
  output$energyplot <- renderPlot({
    # Render a barplot
    ggplot(prod_clean_shiny(),aes(x=TIME_PERIOD, y=Consumed/10**6, fill=Fuel)) + 
              geom_area(alpha=0.6 , size=1, colour="black") + 
            labs(x="Year",
                 y=expression(paste('Energy consumed (',10^6,' GWE)')),
                 title="Cummulative energy production inc. electricity") + 
            theme_bw() +
            scale_fill_brewer(palette = "RdYlBu") + 
            theme(axis.text.x = element_text(angle = 45,hjust=1),panel.grid = element_blank())
  })
}

ui <- fluidPage(    
  # Give the page a title
  titlePanel("Energy production by region"),
  # Generate a row with a sidebar
  sidebarLayout(      
    # Define the sidebar with one input
    sidebarPanel(
      selectInput("region", "Region:", 
                  choices=unique(european_data$Country)),
      hr(),
      helpText("Data obtained from ()")
    ),
    # Create a spot for the barplot
    mainPanel(
      plotOutput("energyplot")  
    )
    
  )
)
shinyApp(ui,server)

```


## Second Plot: 


```{r, message=FALSE, include=FALSE}
## Only electricity - use shiny to extend by country and year::

Total_ele_2020<- elec %>% filter(TIME_PERIOD=="2020-01-01",geo=="DE")
elec_2020 <- data.frame(tapply(Total_ele_2020$OBS_VALUE,list(Total_ele_2020$Fuel), function(x) {sum(x,na.rm=TRUE)}))
summ_elec_1<-data.frame(Consumed=tapply(Total_ele_2020$OBS_VALUE,list(Total_ele_2020$Fuel), function(x) {sum(x,na.rm=TRUE)}))
summ_elec_2020_s<-rownames_to_column(summ_elec_1,"Fuel")

summ_elec_2020_s %>% filter(Fuel!= "Total")%>% arrange(desc(Consumed)) %>% mutate(Fuel=factor(Fuel,levels=Fuel)) %>% ggplot(.,aes(x=Fuel,y=Consumed)) + geom_col() + theme(axis.text.x = element_text(angle = 45,hjust=1))

```

- Plus sort the 5 main energy producers and break down, by year, and structure of energy production

```{r warning=FALSE, message=FALSE, echo=FALSE}

main_energy_geo <- data.frame(rbind(data_prod %>% 
                   filter(Fuel!="Electricity",Fuel!="Total"), elec_years )) %>% 
                   mutate( Fuel= as.factor(Fuel),
                           Fuel = fct_collapse(Fuel,"Other"=c("Non-ren waste","Other","Peat","Heat"))) %>% 
                   filter(TIME_PERIOD>= "2017-01-01",TIME_PERIOD<"2021-01-01",geo %in% c("FR","DE","IT","ES","PL"))%>%
                   group_by(geo,TIME_PERIOD,Fuel) %>%
                   summarise(Consumed = sum(OBS_VALUE)) %>%
                   ungroup(Fuel,TIME_PERIOD) %>%
                   mutate(TIME_PERIOD)%>%
                   arrange(desc(Consumed)) %>% drop_na(Consumed)

main_energy_geo$geo =  with(main_energy_geo, reorder(geo, Consumed, FUN= function(x){1/sum(x)}))

p <- ggplot(main_energy_geo, aes(geo, Consumed/1000, label = geo, fill = Fuel)) + 
     geom_bar(stat="identity") + 
     facet_wrap( ~ format(TIME_PERIOD,"%Y"), scales="free")+
     labs(x="Year and country",
          y=expression(paste('Energy consumed (',10^6,' GWE)')),
          title="Energy production") +
    theme_bw() + 
    scale_y_continuous(limits=c(0,2500))

p + scale_fill_brewer(palette = "RdYlBu") + theme(axis.text.x = element_text(angle = 45,hjust=1))

```


### Ideas:

# 2- Energy used by industry? see predictions + expenditure in the future:


```{r error file, include=FALSE}
#For errors:
## em_sector<- read.csv("Data/emissions_by_sector.csv")

# read.csv("Data/emissions_by_sector.csv") %>% filter(Country_code %in% selection_countries$Country_code)

## em_sector_1 <- em_sector %>% filter(Country_code %in% c("EUA","EUX","EUC")) %>% arrange(Year)
## saveRDS(em_sector_1,"Data/sector_breakdown.rds")

#em_sector <- readRDS("Data/sector_breakdown.rds")
#em_sector_1<- em_sector %>% filter(Sector_name %in% c("1 - Energy","1.A.1 - Energy Industries",
                                                     #"1.A.2 - Manufacturing Industries and Construction",
                 #                                    "1.A.3 - Transport",
                  #                                   "1.A.3.a - Domestic Aviation",
                   #                                  "1.A.3.b - Road Transportation",
                    #                                 "1.A.3.b.i - Cars",
                     #                                "1.A.3.b.ii - Light duty trucks",
                      #                               "1.A.3.b.iii - Heavy duty trucks and buses",
                       #                              "1.A.3.b.iv - Motorcycles",
                        #                             "1.A.3.b.v - Other Road Transportation",
                         #                            "1.A.3.c - Railways",
                          #                           "1.A.3.d - Domestic Navigation",
                           #                          "1.A.3.e - Other Transportation",
                            #                         "1.A.4 - Other Sectors",
                             #                        "1.A.4.a - Commercial/Institutional",
                              #                       "1.A.4.b - Residential",
                               #                      "1.A.4.c - Agriculture/Forestry/Fishing",
                                #                     "1.A.5 - Other Other Sectors",
                                 #                    "2 - Industrial Processes and Product Use",
                                  #                   "2.A - Mineral Industry",
                                   #                  "2.B - Chemical Industry",
                                    #                 "2.C - Metal Industry",
                                     #                "2.D - Non-energy Products from Fuels and Solvent Use",
                                      #               "2.E - Electronics Industry",
                                       #              "2.F - Product Uses as Substitutes for ODS",
                                        #             "3 - Agriculture",
                                         #            "Total emissions (UNFCCC)",
                                          #           "Total emissions with international aviation (EU 2020)"
                                           #          ))

# saveRDS(em_sector_1,"Data/sectors_breakdown.rds")

```


```{r breakdown for sectors (europe wide), warning=FALSE, message=FALSE, include=FALSE}

em_sector_raw<-readRDS("Data/sectors_breakdown.rds")
em_sector<- em_sector_raw %>% select(-c(Format_name,Sector_code,Unit,Notation,PublicationDate,DataSource))
# em_sector_raw%>% select(Unit,Pollutant_name)
unique(em_sector$Pollutant_name)

# Aggregate by years, and summary all emissions:
em_sec_year<-data.frame(Consumed=tapply(em_sector$emissions,list(em_sector$Year), function(x) {sum(x,na.rm=TRUE)}))
em_sec_year<-rownames_to_column(em_sec_year,"Year")

ggplot(em_sec_year,aes(x=Year,y=Consumed))+geom_col()+theme(axis.text.x = element_text(angle = 45,hjust=1))

#Aggregate by sectors:

em_sec_sec<-data.frame(Consumed=tapply(em_sector$emissions,list(em_sector$Sector_name), function(x) {sum(x,na.rm=TRUE)}))
em_sec_sec<-rownames_to_column(em_sec_sec,"Sectors")

ggplot(em_sec_sec,aes(x=Sectors,y=Consumed))+
  geom_col()+
  theme(axis.text.x = element_text(angle = 45,hjust=1),panel.grid = element_blank())


#Main charts to create::


```


```{r warning=FALSE, message=FALSE, echo=FALSE}
# 1.x.(1,2,3,4) - energy production
sec_energy_prod<- em_sector %>% filter(Sector_name %in% c("1.A.1 - Energy Industries",
                                                     "1.A.2 - Manufacturing Industries and Construction",
                                                     "1.A.3 - Transport",
                                                     "1.A.4 - Other Sectors",
                                                     "1.A.5 - Other Other Sectors",
                                                     "2 - Industrial Processes and Product Use",
                                                     "3 - Agriculture")) %>% 
                    select(-c(Parent_sector_code,Pollutant_name)) %>%
                    mutate(Sector_name= as.factor(Sector_name),
                           Sector_name=fct_collapse(Sector_name," Other Sectors"=
                                                      c("1.A.4 - Other Sectors","1.A.5 - Other Other Sectors"))) %>% group_by(Country_code,Year,Sector_name) %>% summarise(Emissions = sum(emissions)) %>% ungroup(Country_code,Year,Sector_name)


levels(sec_energy_prod$Sector_name)=c("Energy industry","Manufacturing & construction","Transport","Other","Industrial processes","Agriculture")

plot_3.1<- ggplot(sec_energy_prod %>% filter(Country_code=="EUA") ,
                  aes(x= as.integer(Year), y=Emissions/1000, group=Sector_name,col=Sector_name)) + 
                  geom_line() + 
                  geom_point(alpha=.5) +  
                  labs(x="Years",
                       y=expression(paste('Emissions (',1000,'s Gigagrams)')),
                       title="Cummulative emissions by sector",
                       col="Sectors") + 
                  theme_bw() + 
                  theme(axis.text.x = element_text(angle = 45,hjust=1), panel.grid = element_blank())+
                  ylim(0,3500)
plot_3.1
```

```{r  warning=FALSE, message=FALSE, echo=FALSE}
#   1.3. (a,bi.ii.iii.iv.v,c,d,e) - modes of transportation (specially if it's seen as main pollutant)
sec_energy_transport<- em_sector %>% filter(Sector_name %in% c("1.A.3.a - Domestic Aviation",
                                                     "1.A.3.b.i - Cars",
                                                     "1.A.3.b.ii - Light duty trucks",
                                                     "1.A.3.b.iii - Heavy duty trucks and buses",
                                                     "1.A.3.b.iv - Motorcycles",
                                                     "1.A.3.b.v - Other Road Transportation",
                                                     "1.A.3.c - Railways",
                                                     "1.A.3.d - Domestic Navigation",
                                                     "1.A.3.e - Other Transportation")) %>% 
                    select(-c(Parent_sector_code,Pollutant_name)) %>%
                    mutate(Sector_name= as.factor(Sector_name),
                           Sector_name=fct_collapse(Sector_name," Other transports"=
                                                      c("1.A.3.b.iv - Motorcycles",
                                                        "1.A.3.b.v - Other Road Transportation",
                                                        "1.A.3.c - Railways",
                                                        "1.A.3.d - Domestic Navigation",
                                                        "1.A.3.a - Domestic Aviation",
                                                        "1.A.3.e - Other Transportation"))
                           ) %>% group_by(Country_code,Year,Sector_name) %>% 
                                 summarise(Emissions = sum(emissions)) %>%  
                                 ungroup(Country_code,Year,Sector_name)

levels(sec_energy_transport$Sector_name)=c("Other transports","Cars","Light duty trucks",
                                            "Heavy duty trucks and buses")

plot_3.2<-ggplot(sec_energy_transport %>% filter(Country_code=="EUA") ,
                 aes(x=as.integer(Year), y=Emissions/1000, group=Sector_name,col=Sector_name)) +
                 geom_line() +
                 geom_point(alpha=.5)+
                 labs(x="Years",
                     y=expression(paste('Emissions (',1000,'s Gigagrams)')),
                     title="European Cummulative emissions by transport", 
                     col="Sectors") + 
                 theme_bw() + 
                 theme(axis.text.x = element_text(angle = 45,hjust=1),
                       panel.grid = element_blank())+
            ylim(0, max(sec_energy_transport$Emissions/1000,na.rm=TRUE)+100)

plot_3.2

```

```{r  warning=FALSE, message=FALSE, echo=FALSE}



#   1.3. (a,bi.ii.iii.iv.v,c,d,e) - modes of transportation (specially if it's seen as main pollutant)
sec_energy_residential<- em_sector %>% filter(Sector_name %in% c("1.A.4.a - Commercial/Institutional",
                                                               "1.A.4.b - Residential",
                                                               "1.A.4.c - Agriculture/Forestry/Fishing",
                                                               "1.A.2 - Manufacturing Industries and Construction",
                                                               "2 - Industrial Processes and Product Use",
                                                               "1.A.3 - Transport",
                                                               "3 - Agriculture")) %>% 
                    select(-c(Parent_sector_code,Pollutant_name)) %>%
                    mutate(Sector_name= as.factor(Sector_name),
                           Sector_name=fct_collapse(Sector_name,"Residential/Commercial"= 
                                                          c("1.A.4.a - Commercial/Institutional","1.A.4.b - Residential")),
                           Sector_name=fct_collapse(Sector_name,"Agriculture"=
                                                          c("1.A.4.c - Agriculture/Forestry/Fishing","3 - Agriculture")),
                           Sector_name=fct_collapse(Sector_name,"Industry"=
                                                          c("1.A.2 - Manufacturing Industries and Construction",
                                                               "2 - Industrial Processes and Product Use"))) %>% group_by(Country_code,Year,Sector_name) %>% summarise(Emissions = sum(emissions)) %>%  
                                 ungroup(Country_code,Year,Sector_name)

levels(sec_energy_residential$Sector_name)=c("Industry","Transport","Residential/Commercial",
                                            "Agriculture")

plot_3.3 <- ggplot(sec_energy_residential %>% filter(Country_code=="EUA") ,
                 aes(x=as.integer(Year), y=Emissions/1000, group=Sector_name,col=Sector_name)) +
                 geom_line() + 
                 geom_point(alpha=.5)+
                 theme_bw() + 
                 theme(axis.text.x = element_text(angle = 45,hjust=1),
                       panel.grid = element_blank()) +
                 labs(x="Years",
                      y=expression(paste('Emissions (',1000,'s Gigagrams)')),
                      title="European Cummulative emissions by sectors",
                      col="Sectors")+
                  ylim(0, 2750)
                       

plot_3.3

```

# 3 - Country-wise pollution and future predictions.

```{r sorting the datasets, warning=FALSE, message=FALSE, echo=FALSE}

#bring the data for the map
world = map_data('world')  %>% mutate(region = ifelse(region=="UK","United Kingdom",region))

europe = c((dictionary_countries %>% 
            mutate(Country=ifelse(Country=="United Kingdom (Convention)","United 
                                  Kingdom",Country),Country_codes=ifelse(Country_codes=="GE","EL",Country_codes)) %>% 
            filter(Country != c("EU-27+UK","EU (KP)","EU-27")))$Country)

europe_codes = c((dictionary_countries %>% 
            mutate(Country=ifelse(Country=="United Kingdom (Convention)","United 
                              Kingdom",Country),Country_codes=ifelse(Country_codes=="GE","EL",Country_codes)) %>% 
              filter(Country != c("EU-27+UK","EU (KP)","EU-27")))$Country_codes)


# Extrapolate the data of the UK - repeat last value:
a<-data_prod %>% mutate(Energy=OBS_VALUE/10^3) %>% 
                filter(Fuel=="Total",TIME_PERIOD=="2020-01-01", geo %in% europe_codes) %>% 
                select(-OBS_VALUE)

b<-data.frame(Fuel="Total",geo="UK",TIME_PERIOD="2020-01-01",Energy = data_prod %>% 
                filter(geo=="UK",TIME_PERIOD=="2019-01-01",Fuel=="Total")%>% 
                mutate(Energy=OBS_VALUE/10^3)%>% 
                select(Energy))

# bring the data for the filling, energy and pollution data:
data_prod_uk<- rbind(a,b)

total_energy<- left_join(data_prod_uk, dictionary_countries %>% 
                           filter(Country != c("EU-27+UK","EU (KP)","EU-27")) %>% 
                           transmute(geo=Country_codes,Country=Country),by="geo") %>% 
                          select(-geo)

fill_pol_en<-left_join(eur_em  %>% 
                          mutate(Country = Geographic_entity, Pollution = eur_em$'2020' ) %>% 
                          filter( Country != "EU-27+UK",Country != "EU (KP)",Country != "EU-27")%>% 
                          select(Country,Pollution),
                  total_energy, by= "Country") %>%   
              transmute(Country,Pollution=Pollution/Energy,region=Country,Fuel,year=TIME_PERIOD,Pol=Pollution,En=Energy)

```


```{r plotting, warning=FALSE, message=FALSE, echo=FALSE}
eur = world[world$region %in% c(fill_pol_en$region,"Iceland","Norway","Turkey","Ukraine","Moldova","Belarus"), ] 
eur <-eur %>% filter(!subregion %in%  c("Jan Mayen","Svalbard"))

clear_plot_background = theme(
  axis.text = element_blank(),
  axis.ticks = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank())

map_plot <- ggplot(data = eur, 
                  mapping = aes(x = long, y = lat, group = group)) +
  coord_map('gilbert') + 
  geom_polygon(colour = 'black', fill = 'white')  + clear_plot_background



eur_em_map = eur %>% 
  left_join(fill_pol_en, by = 'region') %>% drop_na(Country)


map_plot + geom_polygon(data = eur_em_map,  #plot bear population data
               aes(x = long, 
                   y = lat, 
                   group = group, 
                   fill = Pollution), 
               colour = 'white') +
  geom_polygon(colour = 'black', fill = NA) +  #plot country outlines
  scale_fill_viridis(limits = c(0,0.4), breaks = c(0, 0.1, 0.2, 0.3, 0.4),
    guide = guide_colourbar(nbin = 100, draw.ulim = FALSE, draw.llim = FALSE))+
  ft_theme()+ 
  theme(
  axis.text = element_blank(),
  axis.ticks = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank()) +
  labs(x="Years",
       y=expression(paste('Emissions (',1000,'s Gigagrams)')),
       title="European Cummulative emissions by sectors",
       fill=expression(paste("Pollution/Energy (MTC",O[2],"/TWE)")))
  


```
